---
title: "A continuous urbanization index to assess phylogenetic adaptations in urban bird living"


author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
bibliography: references.bib
---

\newpage

# ABSTRACT
Understanding the suite of ecological and life history traits which are associated with urbanization enables a predictive ability to understand the species which are readily suited to adapt with our increasingly urbanizing world. Research investigating this relationship has found contrasting patterns for a specific trait's association with urbanization. Here, we provide a novel and scalable methodology relying on a continuous urbanization index to assess ecological and life history traits which most influence a species' adaptability to persist in urban environments. In a continental-scale analysis, incorporating 477 species, we find that niche breadth, as measured by various types of generalism, and clutch size are among the most important predictors. Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urbanization. Our results suggest that the enhancement and persistence of avian biodiversity in urban envioronments relies on developing suitable habitat to serve a wide range of life history strategies.

*keywords*: birds; brain size; generalists; specialists; life history traits; ecological traits; urbanization index

\newpage
# INTRODUCTION
Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], drastically altering species' composition and ecosystem functions. Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations]. Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation]. However, some species rarely occur within urban areas, despite being relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].

In order to best predict species' responses to future urbanization changes, it is essential to understand what ecological and life history traits are most associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation]. To date, a wide array of traits have been used to predict presence of birds in urban environments. For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a birds' presence in urban environments. Rather than a specific trait that predicts a species' ability to be found in urban environments, it is most likely a combination of traits. For instance, urban birds' presence in Jerusalem was predicted by traits including diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-nesters, and widely distributed [@croci2008does]. Notably, a wide range of results using the aforementioned traits have been reported in the literature, often contrasting one another. Indeed, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains] but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on species' response to urbanization. Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011makes with @moller2009successful & @croci2008does] and niche breadth [cf., @kark2007living with @evans2011makes]. Forming generalizable patterns, at multiple spatial scales [@suarez2018scale], which associate ecological and life history traits with urbanization in birds is critical to help curb the impact urbanization has on avian biodiversity [@sol2017urbanisation].

The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful]. These types of discrete response variables severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011makes]. Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing]. In a notable exception, [@evans2011makes] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.

Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living]. Fortunately, citizen science [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen] provides broad-scale empirical data, allowing us to test hypotheses at spatial and temporal scales previously difficult to attain. These data provide a cost-effective method to research the relationship between species' traits and urbanization in birds. Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global]. Yet, we are unaware of any projects which use citizen science observations to assess species' traits in relation to urban environments over a large spatial and temporal scale.

We use a series of publicly available datasets to implement a novel continuous urbanization index in order to assess ecological and life history traits which most influence a bird's response to urbanization. We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against). Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes]. Our second objective was to develop models which were and were not phylogenetically constrained to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables. These results are especially critical in the future action plans of Australia's birds [@garnett2011action], given Australia is currently undergoing a period of rapid urban growth [@coffee2016visualising].


# METHODS


## Bird observation data
We collated bird observations throughout the whole of continental Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected. Bird observations from outlying islands were eliminated from potential analyses.  

We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new]. For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types]. Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha. Any checklists which were shared among multiple individuals were subsampled to include only one checklist in order to avoid duplicate results. After this sampling, we were left with 637,482 eligible checklists throughout Australia.

Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas under normal circumstances. Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses. This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame.

## Urbanization index
We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat. The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement. There are also a significant number of steps (e.g., filtering out background noise, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs]. For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer. For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols. Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation].

Each checklist's average radiance value was then passed to every bird on that checklist, which provided every species with a distribution of average radiance values (Fig. 1). The median for each species' given distribution of average radiance values served as our response variable in our analyses.

## Biological traits
Predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological]. Potential traits (Appendix 1a) were chosen on the basis of their existing support and hypotheses in the current literature (e.g., [@mcclure1989characterizes; @kark2007living; @evans2011makes]). Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat). Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated. Preliminary analyses including IUCN status (exotic/native) showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analysis to avoid bias. Therefore, we excluded IUCN status as a trait given that the majority of our candidate species were of Least Concern.

Only species which had complete data for the list of candidate traits were considered in the analyses, and we thereby ended with 477 species which met this criteria.

## Statistical analyses
All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].

Before modelling, the predictor variables were investigated for multi-collinearity. We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Supplementary Figure 1), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).

We employed two linear modelling approaches; one with and one without phylogenetic constraints. Our urbanization index was log-transformed in order to meet model assumptions. The predictor variables used in the models were those listed in Appendix 1, with the exception of IUCN status.

We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices. This error was the  proportion of the number of total observations of a species divided by the number of unique localities a species was observed from. This gave more weight in the model-fitting procedure to those species that had more observations in more unique locations. This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices. This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from known roosting sites [e.g., @callaghanavian].

### Non-phylogenetically constrained models
Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables. We then used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes. In order to confirm our linear model results, we employed a model-averaging approach [@grueber2011multimodel] in which we 'dredged' possible subsets of the global model. This was done using the dredge function in the MuMIn package [@barton2016mumin], and we allowed a total of 11 candidate predictors to be fit at any one time in order to avoid over-fitting. This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with $\Delta$AIC < 4. As with the global model, we present the standardized estimates which represent relative effect sizes.

### Phylogenetic analysis

For the phylogenetic portion of our analyses, we used the phylogenetic trees provided by Jetz et al. [@jetz2012global]. This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global]. In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam]. This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.

Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package [@ho2014phylolm]. Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods.  We also  fit the same model to a range of the equally likely trees from [@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.  

### Taxonomic considerations
Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed. The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/]. There were minor differences which existed in common names, scientific names, and recognized species among the three datasets. A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2). Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset. If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher). Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.


# RESULTS
A total of 637,482 checklists resulting in 5,944,819 observations of 580 different species which had > 100 observations were initially investigated for analyses. The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615). The mean number of observations for a species was 10,250. Plots of all 580 species' distributions of average radiance values are presented in Appendix 3. After the 580 candidate species from the eBird dataset were considered based on complete data following both the trait taxonomy and phylogenetic tree taxonomy (e.g., Appendix 2), we were left with 477 species for analysis. A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization index can be found in Appendix 4.

## Urbanization index
For the 477 species included in the models the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (Black-backed Butcherbird) to 11.85 (Powerful Owl; Fig. 2).

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered. Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1). The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion. We generally found weaker signal for behavioral traits compared with morphological traits. For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).

## Model results
### Non-phylogenetic models
The most important predictors of urbanization in the non-phylogenetic global model (Fig. 4; Appendix 5) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032). Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitarily (t = -2.90, p = 0.004).

When this global model was dredged and model-averaged, patterns remained similar (Fig. 4; Appendix 5). Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors. Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001). Urban birds remained significantly negatively associated with feeding solitarily (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).

### Phylogenetically constrained models
The most significant predictors of urbanization in the phylogenetically controlled global model (Fig. 4; Appendix 5) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005). Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.

When the phylogenetic global model was model-averaged, the patterns were similar (Fig. 4; Appendix 5). Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index. Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.

### Differences in non-phylogenetic and phylogenetic models
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the statistical significance of individual predictors (Fig. 4). However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5). Full model results for each of the four analyses can be found in Appendix 5.


# DISCUSSION
Niche breadth, as measured by different types of generalism, was significantly associated with urbanization in our analyses, evidenced by the robustness of this result across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S2). Conversely, specialized feeding strategies were negatively associated with urbanization: granivorous and insectivorous specialists were much less likely to be found in urban environments. These results suggest that urbanization is, in general, excluding specialist species [@sorace2009distribution]. As such, future urban development aimed at increasing avian diversity in urban areas should focus on maintaining a high diversity of habitat [@melles2003urban] both for foraging [@shochat2004linking], and nesting [@fernandez2001habitat], with a focus on intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role]. A species-specific approach to assess particular resources necessary for specialist species may be necessary [@fernandez2004spatial]. Furthermore, a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], which will in turn benefit avian biodiversity in urban environments. 

The non-random distribution of the urbanization index across the phylogeny (Table 1) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that there are phylogenetic constraints which allow certain birds to persist within urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation]. This finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is likely a result of increased statistical power from our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes]. It's not surprising that a larger sample size finds a non-random pattern in urbanization across the bird phylogeny and we demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models. For example, some predictor variables became less crucial in the phylogenetically controlled models compared with the non-phylogenetic models, including some feeding types (i.e., insect specialists) and habitat preferences (i.e., agricultural). For these variables, it is likely that there may be particular clades which include a number of species that are associated with both that particular trait and urbanization [see @felsenstein1985confidence]. Conversely, some of the most important predictor variables (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization when phylogeny was accounted for, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig. S2). Furthermore, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas. While some studies have found a lack of evidence that residual brain size is an important predictor in urban birds [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urbanization in birds [@moller2009successful; @maklakov2011brains; @moller2015brain]. This finding highlights the necessity to investigate the trait-urbanization relationships in a phylogenetically informed manner. 

Certian aspects of the trait-urbanization relationship appear to be consistent, and thus generalizable, across the globe. For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful]. Urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes]. Although we did not find any strong support for our hypotheses on different nesting types, we did find that in general species which tended to be more social were closely associated with urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living]. While these patterns may be generalizable, we also found patterns in our analysis which seem to contrast the global norm. For example, we found that urbanization selects against granivorous species, contrasting the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [cf., @fuller2008garden]. 

As illustrated (above and in Appendix 1), many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies. As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization. Our models assessed the applicability of traits in relation to one another (i.e., we did not run separate models or statistical tests for each specific trait as previously done). The aforementioned differences in our results compared with others is likely due to increased sample size, as well as our use of a continuous measure of urbanization. This analysis was made possible through the use of global broad-scale empirical datasets, which are rapidly expanding our knowledge of urbanization impacts on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen]. This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google]. We used a novel approach, over a broad spatial and temporal scale, to quantify an urbanization index in birds. We did not explicitly account for changes throughout the annual cycle in a species' association with urban areas [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor. We highlight that future studies, especially within migrant-dominated landscapes (e.g., Europe, North America), should account for changes in a species' response to urbanization throughout the full-annual cycle [@hostetler2015full; @marra2015call]. Further, future studies should look to expand this analysis outside the scope of Australia, and also investigate whether these patterns hold true at multiple spatial scales [@suarez2018scale] (i.e., does city size affect the patterns found here; [@sol2017urbanisation; @batary2018non]). Fully understanding the suite of specific species' traits which are associated with urban birds, and placing them in a theoretical landscape ecology context, will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban areas [@evans2011makes].

We provide the largest study to date using a continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments. Our results suggest that generalist species, with large residual brain size, are better prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns]. We highlight that future studies investigating the trait-urbanization relationship should focus on broad-scale data  to form generalizable patterns which rely on continuous measures for improved predictive performance. This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species. In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management. Urban avian biodiversity could be increased by improving the suitability of foraging for granivores and insectivores, especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more]. 


## Acknowledgments
We thank the > 6,000 eBirders who have contributed data in Australia. We also thank the dedicated eBird team at the Cornell Lab of Ornithology and the global network of reviewers. Thanks to Stephen Garnett and Glenn Ehmke for providing us with extent of occurrence estimates for each of the species in our analysis.

\newpage
# REFERENCES

<div id="refs"></div>

\newpage

# Figures

```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance values."}
knitr::include_graphics("finalFigs/Figure1.pdf")
```

```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a log-scale, of the response variable for the 477 species included in the analysis, with five example species highlighted."}
knitr::include_graphics("finalFigs/Figure2.pdf")
```

```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree of the birds with the tips colored based on the urbanization index. The urbanization index is also shown based on the size of the bar associated with a species."}
knitr::include_graphics("finalFigs/Figure3.pdf")
```

```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed in the analysis. A star represents a significant effect. Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set. Full model results can be found in Appendix 5."}
knitr::include_graphics("finalFigs/Figure4.pdf")
```

```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for parameter estimates from the non-phylogenetically constrained and phylogenetically constrained models."}
knitr::include_graphics("finalFigs/Figure5.pdf")
```

\newpage

# Tables
\small
```{r Table 1, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
Table1a <- read_csv("Data/Table1.csv")
colnames(Table1a) <- gsub("X1", "Variable", colnames(Table1a))
is.num <- sapply(Table1a, is.numeric)
Table1a[is.num] <- lapply(Table1a[is.num], round, 3)

Table1b <- read_csv("Data/TableS2.csv")
colnames(Table1b) <- gsub("X1", "Variable", colnames(Table1b))
is.num <- sapply(Table1b, is.numeric)
Table1b[is.num] <- lapply(Table1b[is.num], round, 3)

grouped <- rbind(cbind(Value="Statistic", Table1a), cbind(Value="p-values", Table1b))
tgrp <- xtabs(cbind(Cmean, I, K, K.star, Lambda) ~ Value+Variable, grouped)

table <- ftable(aperm(tgrp, c(3, 1, 2)), row.vars=c("Variable", "Value"))

library(kableExtra)
library(pander)
pander(table, caption = "Results showing the significance of the phylogenetic signal analyses.", split.table=Inf)

```

\newpage

# Supplementary figures
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
     \beginsupplement


```{r Figure S1, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the modelling procedures."}
knitr::include_graphics("finalFigs/FigureS1.pdf")
```

```{r Figure S2, echo=FALSE, fig.cap="Phylogenetic uncertainty for some of the most important predictor variables included in the analysis."}
knitr::include_graphics("finalFigs/FigureS2.pdf")
```

\newpage    

# Supplementary tables
```{r Table S1, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)

collinearity_investigation <- read_csv("Data/TableS1.csv")
collinearity_investigation <- collinearity_investigation[-1,]

kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor variables used in the analysis.", booktabs = T) %>%
kable_styling(latex_options = c("hold_position"))

```
