---
title: "A continental-scale analysis finds generalists, with large clutch size and residual brain mass, are the most urban-tolerant of birds"


author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/ecology-letters.csl
bibliography: references.bib
---

\newpage

# ABSTRACT
Understanding the suite of ecological and life history traits which influence a species' tolerance to urbanization is critical to understand which species are most likely to persist in our increasingly urbanizing world. Previous work investigating these relationships has found a wide array of contrasting patterns for a given trait's association with urbanization. In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we provide a novel and scalable methodology which assesses the ecological and life history traits which most influence a species' adaptability to persist in urban environments. We elucidate generalized, phylogenetically controlled, patterns: bird species which are specialists, with high clutch size, and large residual brain size are among the most urban-tolerant bird species. Conversely, specialized feeding strategies (i.e., insectivores and granivores) are negatively associated with urbanization. Our results suggest that the enhancement and persistence of avian biodiversity in urban environments relies on maintaining and developing a suite of habitats to serve a range of life history strategies.

*keywords*: birds; brain size; generalists; specialists; life history traits; ecological traits; urbanization index

\newpage
# INTRODUCTION
Species specialization is a fundamentally important biological trait which can dictate population-level responses to changing environments. Generally, species with limited niche breadth experience higher extinction risk [@boulangeat2012niche]. Further, certain organisms can more readily adapt their niches in response to changing environments leading to potential ecological inheritance within specific species [@odling2013niche]. The process of urbanization drastically fragments and isolates natural habitats [@marzluff2001restoration; @melles2003urban; @loram2007urban], providing ecologically novel landscapes for the species within. Indeed, some species have successfully adapted to, colonized, and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others remain relatively common in surrounding habitats but rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird]. Urbanization, by selecting for a specific set of species, globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], which is a major concern for conservation [@czech2000economic; @dearborn2010motivations]. Hence, it is important to fully understand the suite of ecological and life history traits most associated with birds that are successful or unsuccessful in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation]. Only then will the species most at-risk of extinction, and thus in need of conservation attention, be best understood and appropriate action subsequently taken (e.g., necessary habitat improvements).

To date, a wide array of ecological and life history traits have been used to predict presence of birds in urban environments. For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate [@kark2007living; @croci2008does; @conole2011functional], diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes], distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have been hypothesized to influence a species' presence in urban environments. For instance, the presence of urban birds in Jerusalem was best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while a regional analysis throughout France and Switzerland demonstrated that urban birds preferred forested habitats, were sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does]. Notably, a wide range of often contrasting results using aforementioned traits have been reported in the literature. For instance, residual brain size was found to be important by Moller and Erritze [-@moller2015brain] and Maklakov *et al.* [-@maklakov2011brains], but Evans *et al.* [-@evans2011makes] and Kark *et al.* [-@kark2007living] found that brain size had no effect on a species' response to urbanization. Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].

The majority of previous studies assign species to discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even binary grouping based on presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful]. These types of discrete response variables severely limits the capacity to generalize results as it assumes species within a group respond to urbanization equally [@evans2011makes]. Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., @kark2007living; @croci2008does; @geschkecompact], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing]. In a notable exception, Evans *et al.* [-@evans2011makes] advocates for a continuous measure of urbanization in which a species is assessed on its entire distribution in response to urban environments.

Furthermore, despite the number of studies which have investigated the relationships between traits and urbanization, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living]. There is a need for broad-scale empirical datasets, which allow us to test hypotheses at large spatiotemporal scales to make generalizable patterns, and citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen]. Indeed, various projects have used citizen science data to elucidate information on abundance and distribution of birds in urban ecosystems [@mccaffrey2005using; @callaghan2017assessing], as well as functional composition of urban birds [@la2014role; @morelli2016evidence; @lepczyk2017global].

The aim of this study was to determine whether there is a suite of life history traits that predicts avian tolerance to urbanization at a continental scale. We hypothesized that a distinct suite of particular traits would be positively associated with urbanization (i.e., urbanization selects for) while a separate suite of traits would be negatively associated with urbanization (i.e., urbanization selects against). For such an analysis to have generality it must (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3), include a broad-range of traits, and (4) employ a continuous measure of response to urbanization. We conducted an analysis that meets these criteria, integrating a series of publicly available datasets to implement a novel continuous urbanization index to assess ecological and life history traits which most influence a bird's response to urbanization. Our first objective was to investigate the presence of a phylogenetic signal in our urbanization index, as previous studies have found a lack of phylogenetic signal in urbanization [@evans2011makes]. Our second objective was to develop models which were and were not phylogenetically controlled to assess species' traits and their association with urbanization, using a suite of continuous and categorical predictor variables.


# METHODS


## Bird observation data
We collated bird observations throughout the whole of continental Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This time-frame was used because this is the period in which eBird data are richest and because this time period minimized undue leverage of mismatch between changes in the response variable measurement (see below) over the given time frame in which bird observations were collected. Bird observations from outlying islands were eliminated from potential analyses.  

We filtered the bird observations [@callaghan2017assessing; @la2014role] by only including observations which were on complete checklists and which followed these protocols: travelling, random, stationary, area, or BirdLife Australia protocols [@barrett2003new]. For further information on eBird protocols see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types. Further, we filtered the checklists to those with a distance travelled of < 5 km or an area covered which was < 500 Ha. Any checklists which were shared among multiple individuals were subsampled to include only one checklist in order to avoid duplicate results. After this sampling, we were left with 637,482 eligible checklists throughout Australia. Seabirds (gannets, petrels, shearwaters, etc.) were omitted from the analyses as we did not expect any correlation with seabirds using urban areas under normal circumstances. Only species which included > 100 observations in the eBird database over the specified time-frame, and that met our criteria were considered for further analyses. This left us with 580 potential species from the eBird dataset which had > 100 observations on eligible checklists in the specified time frame, which was a substantial proportion of the ~730 bird species which are resident in Australia. Given the diverse datasets (see below) used in this analysis, there were differing taxonomic authorities followed (see Appendix 2 for elaboration).

## Urbanization index
We used VIIRS night-time lights [@elvidge2017viirs] as a proxy of a bird's association with urban habitat. The VIIRS night-time lights measures the electric lighting present on Earth's surface which is generally associated with human settlement (Fig. S1). There are also a significant number of steps (e.g., filtering out background noise such as fire, degraded data, and other light source contamination) which are used to ensure the validity of the data and its association with human settlement [@elvidge2017viirs]. For the 637,482 checklists which met our criteria above, we used the checklists' associated spatial coordinates and Google Earth Engine [@gorelick2017google] to assign the average radiance value calculated from the VIIRS layer. For each checklist, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols. Any radiance values which were negative were forced to 0.00001 [*sensu* @ou2015evaluation]. Each checklist's average radiance value was then passed to every bird on that checklist, which provided every species with a distribution of average radiance values (Fig. 1). The median for each species' given distribution of average radiance values served as the response variable in our analyses. Inherently, there is no definitive range for this index, as it solely depends on a species' distribution of observations.

## Biological traits
Many biological and life history traits have been used in previous studies, and potential traits in our study were chosen on the basis of their existing support and hypotheses in the current literature [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes]; we sought to encompass a broad array of traits previously tested. Hence, predictor variables were calculated using a published dataset which contains the biological, ecological, and conservation information for all of Australia's birds [@garnett2015biological]. Table 1 provides a brief overview of the traits that we investigated, while Appendix 1 Appendix 1 provides an overview of all candidate traits, some associated references for each trait, and more detailed information on how the traits were calculated. Traits were parsed in different manners (Appendix 1b), and for any missing value in the dataset, we treated that as evidence against it existing (i.e., if a species had NA for a specific habitat, we assumed that the lack of evidence found by the authors indicates that the species is unlikely to be found in that habitat). Preliminary analyses including IUCN status (exotic/native) showed that exotic birds were strongly associated with urbanization, but given the preferential selection by acclimatization societies for introduction of species with proven synanthropy [@blackburn2009avian], exotic species were omitted from the final analysis to avoid bias. Therefore, we excluded IUCN status as a trait because the majority of our candidate species were of Least Concern. Only species which had complete data for the list of candidate traits were considered in the analyses, and 477 species met this criteria.

## Statistical analyses
All models and analyses were conducted in R [@r2017coreteam] and relied heavily on the tidyverse workflow [@wickham2017tidyverse].

Before modelling, the predictor variables were investigated for multi-collinearity. We used the corrplot package [@corrplot2017] to show correlation among continuous variables (Fig. S2), and used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation amongst each other (Supplementary Table 1).

We employed two linear modelling approaches; one with and one without phylogenetic constraints. Our urbanization index was log-transformed in order to meet model assumptions. The predictor variables used in the models were those listed in Table 1.

We weighted the linear regression models [@solon2015weighting] using a source of error associated with our urbanization indices. This error was the proportion of the number of total observations of a species divided by the number of unique localities a species was observed from. This gave more weight in the model-fitting procedure to those species that had more observations in more unique locations. This was done based on the assumption that those species which had the most observations would have the most reliable urbanization indices. This approach also addresses biases in sampling of eBird data in which some birds can have a high number of observations, but from a small number of locations such as known roosting sites [e.g., @callaghan2018avian].

### Non-phylogenetically controlled models
Given that we hypothesized any of the potential candidate traits could significantly drive the response in our urbanization index, we first ran one global model including all 22 predictor variables. We used the standardize function from the arm package [@gelman2008scaling; @gelman2016arm] to standardize model estimates to represent relative effect sizes. In order to confirm our linear model results, we employed a model-averaging approach [@grueber2011multimodel] in which we 'dredged' possible subsets of the global model. This was done using the dredge function in the MuMIn package [@barton2016mumin], and we allowed a total of 11 candidate predictors to be fit at any one time in order to avoid over-fitting. This left a total of 784,626 models which were assessed, and the model-averaged parameter estimates were calculated from the top models of those with $\Delta$AIC < 4. As with the global model, we present the standardized estimates which represent relative effect sizes.

### Phylogenetic analysis
As the correlation between traits and urbanization may be influenced by the shared evolutionary history of the species [@felsenstein1985phylogenies], we also used phylogenetically informed models to test our hypotheses. Phylogenetically informed models are descendants of phylogenetic independent contrasts [@felsenstein1985phylogenies], which in simple terms, tests for repeated evolutionary association between the traits of interest. Or in this particular case, whether the traits of interest (e.g., large brain size at a given body size) repeatedly allow for the use of urban environments wherever it evolved across the bird phylogeny. We used the phylogenetic trees provided by Jetz et al. [@jetz2012global]. This phylogeny represents the first completely mapped phylogeny of 9,993 birds [@jetz2012global]. In order to assess whether there was a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] of the urbanization index and the continuous predictor variables used in our study, we used the phylosignal package [@keck2016phylosignal] in the R computing environment [@r2017coreteam]. This was also conducted for each of the continuous response variables in order to assess the relative contribution of phylogeny on the predictor variables.

Phylogenetic models were fit based on the same global model as above, but using the 'phylolm' function from the phylolm package which fits a linear regression model where the likelihood is linear in the number of tips in the tree [@ho2014phylolm]. Similar to above, we model-averaged the global phylogenetic model using the aforementioned methods. To account for aspects of phylogenetic uncertainty, we also fit the same model to a range of equally likely trees from Jetz *et al.* [-@jetz2012global] examining the distribution of putatively important parameters across the distribution of trees.  


# RESULTS
A total of 316,306 checklists resulting in 5,944,819 observations of 477 different species were included in analyses. The least often recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104) while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615). The mean number of observations for a species was 10,250. Plots of all 477 species' distributions of average radiance values are presented in Appendix 3. A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization indices can be found in Appendix 4.

## Urbanization index
For the 477 species included in the analysis the mean $\pm$ SD urbanization index was 0.91 $\pm$ 1.76, with a range from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig. 2).

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Fig. S3; Table 2), as it was significantly different than random for 4 out of the 5 indices considered. For example, note the distinct signal located throughout the waterbirds' clades in Figure 3 or the significantly large index for *Manorina melanocephala*, Noisy Miner, compared to the rest of the honeyeaters. Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 2). The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion. We generally found weaker signal for behavioral traits compared with morphological traits. For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 2).

## Model results
### Non-phylogenetic models
The most important predictors of urbanization (i.e., the largest parameter estimates - shown in parantheses) in the non-phylogenetic global model (Fig. 4; Appendix 5) were feeding habitat generalism (0.79), breeding habitat generalism (0.76), and clutch size (0.73). Further, significance was found for feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban-tolerant birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but less so than the previous traits. Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), insectivorous diet (t = -3.96, p < 0.001), and solitary feeding (t = -2.90, p = 0.004). The strongest negative responses, in rank of their parameter estimates, were solitary feeding (-0.92), insectivorous diet (-0.85), grassland/shrubland habitat association (-0.733), and tree/forest habitat association (-0.63).

When this global model was dredged and model-averaged to investigate for specific suites of traits, patterns remained similar (Fig. 4; Appendix 5). However, the most important variable, based on standardized parameter estimates was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28). Breeding habitat (0.15) and feeding habitat (0.12) generalism parameter estimates were substantially smaller than in the non model-averaged approach. Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors. Additionally, birds which were classified in the traits database as having strong association with agricultural habitats were significantly associated with urbanization (z = 3.40, p = 0.001). Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001). Solitary feeding (-1.00) remained the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).

### Phylogenetically controlled models
Brain residual size (1.25) had the largest standardized parameter estimate for the phylogenetically controlled global model (Fig. 4; Appendix 5), followed by feeding habitat generalism (1.18), clutch size (0.91), and body size (0.82).The most significant predictors of urbanization in the phylogenetically controlled global model  were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism (t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005). Nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.

Similarly, upon model-averaging (Fig. 4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) showed the strongest association with urbanization. Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were all statistically significant. Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization, with no known nest aggregation (-1.22), and species associated with grasslands/shrublands (-1.16) showing the strongest negative response.

### Differences in non-phylogenetic and phylogenetic models
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates (Appendix 5), and the statistical significance of individual predictors (Appendix 5; Fig. 4). However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5), showing the strongest relative response in the phylogenetic global model. Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Appendix 5; Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5). Full model results for each of the four analyses can be found in Appendix 5.


# DISCUSSION
The most urban-tolerant species in our analyses (i.e., the ones with a higher urbanization index) had larger niche breadths, as measured by different forms of generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) were less likely to be found in urban environments. These results were robust across non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S4). These results suggest that the process of urbanization is, in general, operating to exclude specialist species [@sorace2009distribution]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent with the observation of reduced phylogenetic diversity in urban areas [@la2018phylogenetic], suggesting that particular phylogenetic constraints allow certain species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

Our finding of a strong phylogenetic signal contrasts with previous studies [@evans2011makes], but this is unsurprising given our larger sample size: 477 species in this study compared with 55 species in Evans *et al.* [-@evans2011makes]. We further demonstrated the importance of treating species as non-independent by considering phylogeny in the context of the linear models, and thus highlight the importance of investigating trait-urbanization relationships in a phylogenetically informed manner. For example, some of the most important predictor variables in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) showed a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models, and this affect was robust relative to a consideration of phylogenetic uncertainty (Fig. S4); and other traits showed the opposite trend. Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models, suggesting that within a clade the species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] are most suited to surviving in urban areas. While some studies have found a lack of evidence that residual brain size is an important predictor in urban-tolerant species [@kark2007living; @evans2011makes], our results suggest that residual brain size is indeed heavily correlated with urban-tolerance [@moller2009successful; @maklakov2011brains; @moller2015brain]. 

Many traits have been used to identify a species' adaptability to urban environments, but results have frequently differed among studies. As previously noted, a major shortcoming of previous studies has been the focus on discrete classifications of species based on their responses to urbanization. Our models assessed the applicability of traits in relation to one another [i.e., we did not run separate models or statistical tests for each specific trait as previously done; e.g., @kark2007living] thus allowing comparable effect sizes. Despite this, there appear to be some consistent, and thus generalizable, trends within the trait-urbanization complex across the globe. For instance, our finding of the relative importance of annual fecundity is consistent with work in Europe [@croci2008does; @moller2009successful]. Additionally, urbanization appears to generally select against insectivorous species [@chace2006urban; @evans2011makes]. Although we did not find any strong support for our hypotheses on different nesting types, we found a positive association between the degree of sociality and urbanization, consistent with other findings [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living]. While the aforementioned patterns may be generalizable, we also found patterns in our analyses which seem to contrast the global norm. For example, we found that urbanization selects against granivorous species, contrasting with the general trend [@chace2006urban; @kark2007living; @evans2011makes], but this could be a result of relatively little supplementary feeding in Australian cities [compared to European cities for instance; @fuller2008garden].

Our analysis was made possible through integration of broad-scale empirical datasets, which are rapidly expanding our knowledge of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen]. This is largely due to the growing use of citizen science data [@silvertown2009new] in combination with advances in remote sensing technology [@gorelick2017google]. We did not explicitly account for changes throughout the annual cycle in a species' association with urban environments [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered as a predictor. We emphasize that future studies, especially within migrant-dominated systems (e.g., Europe, North America), should account for changes in a species' urbanization index throughout the full-annual cycle [@hostetler2015full; @marra2015call]. This work should be further extended to test whether our findings can accurately predict species which are commonly found in urban environments in other parts of the world, based on a known suite of traits associated with a given set of species.

Fully understanding the suite of traits which are associated with urban-tolerant species will ultimately allow for a better understanding of how to best preserve and maintain avian biodiversity in urban environments [@evans2011makes]. Given specialist species are known to have strong negative responses to landscape fragmentation and disturbance [@devictor2008distribution], sustainable urban development with a goal of maintaining biological diversity in urban areas should focus on providing intact vegetation [@sushinsky2013should] connected by corridors [@litteral2017role]. We recommend that a focal point of avian biodiversity conservation should be on increasing insect abundance and diversity in urban environments [@baldock2015uk], especially important given the recent noted collapse of insects in other parts of the world [@hallmann2017more]. A species-specific approach, which assesses particular resources necessary for specialist species, for example focusing on foraging [@shochat2004linking] and nesting [@fernandez2001habitat] requirements, is likely necessary. To this end, our urbanization index can be used as a potential means to provide quantifiable targets for local restoration projects. For example, a particular restoration project should aim to add species which have lower urbanization indices into their local avifauna, but we note that these indices may differ at different spatial scales [@suarez2018scale]. 

We provide a novel continuous urbanization index to assess ecological and life history traits which are associated with a species' adaptability to urban environments. Future studies investigating the trait-urbanization relationship should focus on broad-scale data to form generalizable patterns which rely on continuous measures for improved predictive performance. Our results suggest that generalist species, with high clutch sizes and large residual brain sizes, are best prepared to forage and survive in the ecologically novel aspects of urban environments [@shochat2006patterns]. In the context of our specific findings, Australia is a relatively young country with rapid urbanization currently underway [@coffee2016visualising] and fully understanding how birds respond to urban development is critical to inform future management. Urban avian biodiversity could be increased by emphasizing the critical importance of intact vegetation [@sushinsky2013should] for specialist species [@devictor2008distribution], and concomitantly improving the suitability of foraging for granivores and insectivores.


## Acknowledgments
We thank the > 6,000 eBirders who have contributed data in Australia. We also thank the dedicated eBird team at the Cornell Lab of Ornithology and the global network of reviewers. Thanks to Stephen Garnett and Glenn Ehmke for providing us with extent of occurrence estimates for each of the species in our analysis. Thanks to Corrine Edwards for providing the illustrations used in Fig. 3.

\newpage
# REFERENCES

<div id="refs"></div>

\newpage

# Figures

```{r Figure 1, echo=FALSE, fig.cap="An example of five different species' average radiance value distributions, on a logarithmic scale. These species were chosen because they represent species with different distributional responses."}
knitr::include_graphics("finalFigs/Figure1.pdf")
```


```{r Figure 2, echo=FALSE, fig.cap="The distribution, on a logarithmic scale, of the urbanization index for the 477 species included in the analysis, with five example species highlighted. The urbanization index was taken as the median of the distribution for a species' average radiance values, as demonstrated in Figure 1."}
knitr::include_graphics("finalFigs/Figure2.pdf")
```


```{r Figure 3, echo=FALSE, fig.cap="A phylogenetic tree, based on Jetz et al. (2012) of the 477 bird species in the analysis, demonstrating the clustering of the urbanization index in regards to phylogeny. The tree demonstrates two representations of the urbanization index for each species: (1) coloring of the tips where dark red equates to a low urbanization index and dark blue equates to a high urbanization index, and (2) the size of the barplot at a tip shows the relative magnitude of the urbanization index. The urbanization index was taken as the median of the distribution for a species' average radiance values. Figure S3 shows a reference phylogenetic tree with the names of each tip displayed, and this link (https://johnwilshire.github.io/bird_urbanness_web_app/) provides an interactive visualization in order to easily identify each of the 477 species in the tree and their respective urbanization indices. For a full list of the 477 species included and their urbanization index see Appendix 4."}
knitr::include_graphics("finalFigs/Figure3.pdf")
```


```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed in the analysis. A star represents a significant effect. Note that the model-averaged results do not contain every possible predictor variable in the case that a given predictor variable was not included in the top model set which was model-averaged. Full model results can be found in Appendix 5."}
knitr::include_graphics("finalFigs/Figure4.pdf")
```


```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for 22 standardized parameter estimates from the non-phylogenetically controlled and phylogenetically controlled global models, as modelled based on 477 species."}
knitr::include_graphics("finalFigs/Figure5.pdf")
```


\newpage

# Tables
\small
```{r Table 2, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
Table1a <- read_csv("Data/Table1.csv")
colnames(Table1a) <- gsub("X1", "Variable", colnames(Table1a))
is.num <- sapply(Table1a, is.numeric)
Table1a[is.num] <- lapply(Table1a[is.num], round, 3)

Table1b <- read_csv("Data/TableS2.csv")
colnames(Table1b) <- gsub("X1", "Variable", colnames(Table1b))
is.num <- sapply(Table1b, is.numeric)
Table1b[is.num] <- lapply(Table1b[is.num], round, 3)

grouped <- rbind(cbind(Value="Statistic", Table1a), cbind(Value="p-values", Table1b))
tgrp <- xtabs(cbind(Cmean, I, K, K.star, Lambda) ~ Value+Variable, grouped)

table <- ftable(aperm(tgrp, c(3, 1, 2)), row.vars=c("Variable", "Value"))

library(kableExtra)
library(pander)
pander(table, caption = "Results showing the  phylogenetic signal for both predictor and response traits.  Because there are a number of indices in the literature, with specific strengths and weaknesses, we present here all available indices (Keck et al. 2016).", split.table=Inf)

```

\newpage

# Supplementary figures
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
     \beginsupplement

```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane (top right), Melbourne (bottom left), and Sydney (bottom right). Values are measured as average radiance, shown in red. For each checklist in our analysis, the average radiance was calculated within a 5 km buffer, in order to match the spatial scale of the checklists included, and this led to the distributions shown in Figure 1."}
knitr::include_graphics("finalFigs/FigureS1.pdf")
```

```{r Figure S2, echo=FALSE, fig.cap="A correlation plot for the continuous variables used in the analysis."}
knitr::include_graphics("finalFigs/FigureS2.pdf")
```

```{r Figure S3, echo=FALSE, fig.cap="A reference phylogenetic tree, with labeled tips, based on Jetz et al. (2012) of the 477 bird species in the analysis. Appendix 4 lists the 477 species."}
knitr::include_graphics("finalFigs/FigureS3.pdf")
```

```{r Figure S4, echo=FALSE, fig.cap="To account for one aspect of phylogenetic uncertainty, we fits the same phylolm model to 1000 of the family of trees presented by Jetz et al. (2012).  The density plots are the estimated parameters for most important variables across models fit to the collection trees from Jetz et al. (2012)"}
knitr::include_graphics("finalFigs/FigureS4.pdf")
```

\newpage    

# Supplementary tables
```{r Table S1, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)

collinearity_investigation <- read_csv("Data/TableS1.csv")
collinearity_investigation <- collinearity_investigation[-1,]

kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor variables used in the analysis.", booktabs = T) %>%
kable_styling(latex_options = c("hold_position"))

```
