---
title: "Generalists are the most urban-tolerant of birds: an analysis of ecological and life history traits using a novel continuous urbanization index"


author: |
  | Corey T. Callaghan^[Corresponding author: c.callaghan@unsw.edu.au] $^1$, Richard E. Major $^1$$^,$$^2$, John H. Wilshire $^1$, John M. Martin $^1$$^,$$^3$, Richard T. Kingsford $^1$, William K. Cornwell $^1$$^,$$^4$
  | $^1$Centre for Ecosystem Science, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia; $^2$Australian Museum Research Institute, Sydney, NSW, Australia; $^3$Royal Botanic Gardens and Domain Trust, Sydney, NSW, Australia; $^4$Evolution and Ecology Research Centre, School of Biological, Earth and Environmental Sciences, UNSW Sydney, Sydney, NSW, Australia
date: ""
output:
  rmarkdown::pdf_document:
    fig_caption: yes    
    includes:  
      in_header: preamble-latex.tex
fontsize: 12pt
link-citations: no
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/proceedings-of-the-royal-society-b.csl
bibliography: references.bib
---

\newpage

# ABSTRACT
Identifying which ecological and life history traits influence a species' tolerance to urbanization is critical to understanding the trajectory of biodiversity in an increasingly urbanizing world. There is evidence for a wide array of contrasting patterns for single trait associations with urbanization. In a continental-scale analysis, incorporating 477 species and > 5,000,000 bird observations, we developed a novel and scalable methodology that evaluated the ecological and life history traits which most influence a species' adaptability to persist in urban environments. We identified generalized, phylogenetically controlled patterns: bird species which are generalists (i.e., large niche breadth), with high clutch size, and large residual brain size are among the most urban-tolerant bird species. Conversely, specialized feeding strategies (i.e., insectivores and granivores) were negatively associated with urbanization. Enhancement and persistence of avian biodiversity in urban environments probably relies on protecting, maintaining, and restoring diverse habitats serving a range of life history strategies.

*Keywords*: eBird, citizen science, phylogenetics, life history traits, avian, brain size

\newpage
# INTRODUCTION
Species' specialization, based on a suite of biological traits, can dictate population-level responses to changing environments. Generally, species with limited niche breadth experience higher extinction risk than those with wider niche breadth [@boulangeat2012niche]. Further, some organisms have traits that allow for adaptation of their niches with changing environments, leading to potential ecological inheritance [@odling2013niche]. Urbanization drastically changes natural habitats through destruction, fragmentation, and isolation [@marzluff2001restoration; @melles2003urban; @loram2007urban], often providing ecologically novel landscapes for species. Some species successfully adapt, colonize, and thrive in urban environments [@mckinney2002urbanization; @chace2006urban; @evans2009independent; @lowry2013behavioural; @mcdonnell2015adaptation], while others, common in surrounding natural habitats, rarely occur in urban environments [@tratalos2007bird; @evans2011makes; @gatesire2014bird]. Urbanization selects for specific sets of species, globally threatening biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects], a major concern for conservation [@czech2000economic; @dearborn2010motivations]. It is important to identify which ecological and life history traits most associate with successful or unsuccessful species in urban environments [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful; @mcdonnell2015adaptation]. This can improve understanding of relationships between biodiversity and urbanization, and guide protection of the species most at-risk of extinction, through habitat protection, management, and restoration.

Many ecological and life history traits are associated with bird species in urban environments, including degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution]; nesting substrate [@kark2007living; @croci2008does; @conole2011functional]; diet [@beissinger1982effects; @fuller2008garden; @major2010museum; @evans2011makes]; range size [@jokimaki2000artificial; @chace2006urban; @croci2008does]; migratory status [@friesen1995effects]; niche breadth [@clergeau2006avifauna; @kark2007living; @evans2011makes]; and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific] (see Appendix 1). In Jerusalem, urban bird species were best predicted by diet, degree of sociality, sedentariness, and preferred nesting sites [@kark2007living], while, throughout France and Switzerland, urban bird species were positively associated with forested habitats, sedentary, omnivorous, high-elevation nesters, and widely distributed [@croci2008does]. There are also contrasting results: residual brain size in birds is sometimes positively associated with urbanization [@maklakov2011brains; @moller2015brain] and sometimes not [@kark2007living; @evans2011makes]. Similarly, contrasting results have been reported for annual fecundity [cf., @croci2008does; @moller2009successful; @evans2011makes] and niche breadth [cf., @kark2007living; @evans2011makes].

Most analyses assign species to discrete a priori groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or strictly presence/absence within urban environments [e.g., @bonier2007urban; @moller2009successful]. These approaches severely limit the capacity to generalize, as it assumes grouped species respond equally to urbanization [@evans2011makes]. Terminology also varies among studies [cf., @kark2007living; @croci2008does; @geschkecompact], affecting interpretation [@mcdonnell2015adaptation; @fischer2015categorizing]. There is clearly a need for continuous measures of urbanization [@evans2011makes], where a species is assessed over its entire distribution in response to urban environments.

Most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does] or temporally limited to a single breeding season [@blair1996land] or even two-month survey period [@kark2007living]. Broad-scale empirical data allow hypotheses to be tested at large spatiotemporal scales. Citizen science can provide such data [see reviews by @devictor2010beyond; @tulloch2013realising; @bonney2014next; @kobori2016citizen], and can be used to describe patterns of abundance, distribution, and functional composition of birds in urban ecosystems [@mccaffrey2005using; @la2014role; @morelli2016evidence; @lepczyk2017global @callaghan2017assessing].

We aimed to determine what suite of life history traits predicted avian tolerance to urbanization at a continental scale. We predicted that different traits would positively and negatively associate with urbanization. For generality, our analyses had to (1) be comprehensive in faunal coverage, (2) account for phylogenetic constraints, (3) include a broad-range of traits, and (4) employ a continuous measure of urbanization. We integrated a series of large datasets to establish a novel continuous urbanization index to test associations with ecological and life history traits influencing species' responses to urbanization. We first investigated whether our urbanization index was statistically independent in relation to phylogeny (i.e., phylogenetic signal - [@revell2008phylogenetic; @evans2011makes]), given our comprehensive sample size. We then developed models, including and excluding phylogenetic effects, to test for associations between our urbanization index and a suite of continuous and categorical predictor variables representing species' traits.


# METHODS
## Bird observation data
We collated bird observations throughout mainland Australia from eBird [@sullivan2009ebird; @sullivan2014ebird], a large-scale empirical dataset contributed by citizen scientists. Volunteer birdwatchers submit lists of species seen or heard at a given location, over a user-determined duration and survey area [@wood2011ebird], often comparable to trained observers [@callaghan2015efficacy]. We downloaded the eBird basic dataset (version ebd_relFeb-2018; available here: https://ebird.org/data/download), and used all observations between January 1, 2010 - February 28, 2018. This corresponded to the richest period of eBird data, minimizing undue leverage of mismatch between changes in response variables over which bird observations were collected.

We filtered bird observations [@callaghan2017assessing; @la2014role], only including observations on complete checklists and which were either travelling, random, stationary, area, or followed BirdLife Australia protocols [@barrett2003new] (see here: http://help.ebird.org/customer/portal/articles/1006209-understanding-observation-types). We then filtered the checklists to those which travelled < 5 km or covered < 500 Ha, to ensure the highest quality data were used. Any checklists shared among multiple observers were subsampled, including only one checklist to avoid duplication. This provided 637,482 eligible checklists throughout Australia. Seabirds (e.g., gannets, petrels, shearwaters) were omitted from analyses as we did not expect any correlation with seabirds using urban areas. Further, only species with > 100 observations were analyzed. In total 580 species, most (~80%) of Australia's bird species, were candidates for analyses. We used differing taxonomic authorities (see Appendix 2).

## Urbanization index
We used VIIRS night-time lights [@elvidge2017viirs] as a continuous measure for urban habitat, measuring the electric lighting associated with human settlement (Fig. S1). This has filtered out background noise (e.g., fires, degraded data, and other light source contamination) to ensure validity of the data and its association with human settlement [@elvidge2017viirs]. For our checklists and their associated species, we used the spatial coordinates and Google Earth Engine [@gorelick2017google], to assign the average radiance value, calculated from the VIIRS layer. For each checklist, the average radiance was calculated within a 5 km buffer, to match the spatial scale of the checklists included and to minimize any potential bias in eBird sampling protocols. Any negative radiance values were forced to 0.00001 [*sensu* @ou2015evaluation]. Every bird species subsequently had a distribution of average radiance values (Fig. 1), with the median of each species' distribution becoming the response variable. There was no definitive range for this index, dependent on a species' distribution of observations.

## Ecological and life history traits
Ecological and life history traits were chosen on the basis of existing support [e.g., @mcclure1989characterizes; @kark2007living; @evans2011makes] (Table 1; Appendix 1), including a wide array of potential traits for all of Australia's birds [@garnett2015biological]. A total of 22 traits, grouped into 7 categories, were included (Table 1): life history traits (n=3), measures of niche breadth (n=4), specific habitat preferences (n=3), specific nesting preferences (n=2), specific diet preferences (n=4), movement and range size (n=3), and degree of sociality (n=3). For any missing trait in the dataset [@garnett2015biological] for a species, we assumed it did not exist. For example, if a species had no data for a specific habitat, it was deemed not to occur in that habitat. Our preliminary analyses, which included IUCN status, showed that exotic birds strongly associated with urbanization; these were omitted from the final analyses to avoid bias, given the preferential selection by acclimatization societies for their proven synanthropy [@blackburn2009avian]. We also excluded IUCN status because most (96%) of our candidate species were of Least Concern. In total, 477 species had complete data for all candidate traits and were included in analyses.

## Statistical analyses
Before modelling of predictor variables (Table 1), we investigated multi-collinearity among continuous variables (Fig. S2), using the corrplot package [@corrplot2017]. We used generalized variance inflation factors [@fox1992generalized], from the car package [@fox2011companion], which accounts for the degrees of freedom for a particular term to show minimal correlation among each other; all traits < 2 suggested negligible correlation (Supplementary Table 1). We used two independent linear modelling approaches: one without and one with phylogenetic effects. Phylogenetic effects are defined as the tendency of closely related organisms to have similar life history components due to their shared evolutionary history [@derrickson1988taxon; @mckitrick1993phylogenetic]. Both approaches were used because of shortcomings [@cadotte2017phylogenies] and importance [@losos2008phylogenetic; @graham2018phylogenetic] of incorporating evolutionary distinctness in ecological research.

Linear regression models were weighted [@solon2015weighting], using a source of error associated with our urbanization indices, which were log-transformed to meet model assumptions. This error was the proportion of the number of total observations for a species, divided by the number of its unique localities (Fig. S2 shows the correlation between the two). This weighted the model-fitting to those species with more observations in more unique locations, likely to be the most reliable urbanization indices. This approach also addressed biases in sampling of eBird data, where some birds have a high number of observations from a small number of locations such as known roosting sites [e.g., @callaghan2018avian]. All models and analyses were conducted in R [@r2017coreteam], relying on the tidyverse workflow [@wickham2017tidyverse]. Statistical significance was concluded as $\alpha$ < 0.05. Variability of means is reported as standard deviations.

## Non-phylogenetically controlled models
We first ran a global model, including all 22 predictor traits, standardizing model estimates to represent relative effect sizes, using the standardize function from the arm package [@gelman2008scaling; @gelman2016arm]. To corroborate our global model results, we employed a model-averaging approach [@grueber2011multimodel], 'dredging' all possible subsets of the global model with the dredge function in the MuMIn package [@barton2016mumin]. We fitted 11 candidate traits at any one time to avoid over-fitting, assessing 784,626 models. Model-averaged parameter estimates were calculated from the subsequent top models, with $\Delta$AIC < 4, where standardized estimates represented relative effect sizes of respective traits.

## Phylogenetic analysis
### Phylogenetic signal
As associations between traits and urbanization may be influenced by shared evolutionary history of a species [@felsenstein1985phylogenies], we also modelled this phylogenetically. First, we tested for a phylogenetic signal [@revell2008phylogenetic; @munkemuller2012measure] associated with our urbanization index and for each of the continuous predictor traits (Table 1), using published phylogenetic trees [@jetz2012global]. This was defined as the statistical nonindependence among species' traits, due to their phylogenetic relatedness [@revell2008phylogenetic], assessing relative contribution of phylogeny on a given trait. There are different indices used to assess phylogenetic signal, each with strengths and weaknesses [@keck2016phylosignal]; we assessed all available indices (Blomberg's K and K*, Abouheif's Cmean, Moran's I, and Pagel's Lambda), using the phylosignal package [@keck2016phylosignal] in R.

### Phylogenetically controlled models
Second, we used phylogenetically informed models to test for repeated evolutionary associations between traits of interest [@felsenstein1985phylogenies]. Phylogenetic models were fitted based on the same global model as the non-phylogenetic models, but using the 'phylolm' function [@ho2014phylolm], which fits a linear regression model where the likelihood is linear in the number of tips in the tree. We similarly model-averaged the global phylogenetic model. To account for phylogenetic uncertainty, we fitted the same model to 1000 highly probable trees from Jetz *et al.* [-@jetz2012global], examining the distribution of putatively important traits across the distribution of trees.  


# RESULTS
We analyzed 316,306 checklists, including 5,944,819 observations of our 477 different species. The least recorded species was *Turnix melanogaster* (Black-breasted Buttonquail; N=104), while the most recorded species was *Cracticus tibicen* (Australian Magpie; N=158,615). The mean number of observations for a species was 10,930 $\pm$ 19,498. Distributions of measured VIIRS average radiance values (shown for each species in Appendix 3), traits, number of observations, and urbanization indices for the 477 species varied considerably (Appendix 4). The mean urbanization index was 0.91 $\pm$ 1.76, ranging from 0.008 (*Cracticus mentalis* Black-backed Butcherbird) to 11.85 (*Ninox strenua* Powerful Owl; Fig. 2).

## Non-phylogenetic models
The most important predictors of urbanization (i.e., the largest parameter estimates - shown in parentheses) in the non-phylogenetic global model (Fig. 4; Appendix 5) were feeding habitat generalism (0.79), breeding habitat generalism (0.76), and clutch size (0.73). Further, feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006) were significant. Urban-tolerant birds were also significantly positively associated with agricultural habitats (t = 2.88, p = 0.004) and more likely to be migratory (t = 2.15, p = 0.032), but had smaller effect sizes than previous traits. Conversely, urban-tolerant birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), group foraging (t = -2.24, p =0.025), granivorous diet (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), and grass/shrubland habitats (t = -4.26, p < 0.001), with insectivorous diet (t = -3.96, p < 0.001) and solitary feeding (t = -2.90, p = 0.004). The strongest negative associations with the urbanization index were solitary feeding (-0.92), insectivorous diet (-0.85), grass/shrubland habitat association (-0.733), and tree/forest habitat association (-0.63).

With model-averaging, patterns remained similar (Fig. 4; Appendix 5). However, the most strongly associated trait was clutch size (0.75), followed by birds associated with agricultural habitats (0.58), and then diet generalism (0.28). Breeding habitat (0.15) and feeding habitat (0.12) generalism parameter estimates were substantially smaller than in the non model-averaged approach. Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) were all significantly associated with our urbanization index. Additionally, species associated with agricultural habitats were significantly associated with urbanization (z = 3.40, p = 0.001). Urban-tolerant birds remained significantly negatively associated with solitary feeding (z = 3.39, p = 0.001),  insectivorous diet (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), group foraging (z = 3.16, p = 0.002), granivorous diet (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001). Solitary feeding (-1.00) was the strongest negatively associated trait with urbanization, followed by insectivorous diet (-0.98).

## Phylogenetic analyses
### Phylogenetic signal
There was a distinct relationship between the phylogenetic signal and our urbanization index (Fig. 3; Fig. S3; Table 2), significantly different from random for 4 out of 5 indices (Cmean, I, K.star, and Lambda). For example, the urbanization index was disproportionately important with respect to the phylogenetic tree (Fig. 3), evidenced by the highest urbanization indices clustered throughout the waterbird clade. Further, within a specific clade, there were some species with substantially higher urbanization indices (e.g., *Manorina melanocephala*, Noisy Miner, cf. honeyeaters). We generally found weaker signals for behavioral, compared with morphological traits. For behavioural traits, the estimates of phylogenetic signal were weaker than Brownian motion, but usually significant (Table 2). Additionally, there was a strong phylogenetic signal for body size and brain residual size (Table 2).

### Phylogenetically controlled models
Brain residual size (1.25) had the largest standardized parameter estimate for the phylogenetically controlled global model (Fig. 4; Appendix 5), followed by feeding habitat generalism (1.18), clutch size (0.91), and body size (0.82).The most significantly positive traits associated with urbanization were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism (t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005). Conversely, nest generalism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding (t = -3.25, p = 0.001), and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with urbanization.

With model-averaging (Fig. 4; Appendix 5), feeding habitat generalism (1.30), clutch size (1.15), body size (0.66), and residual brain size (0.32) were most strongly associated with our urbanization index, while species with no known nest aggregation (-1.22) and species associated with grass/shrubland (-1.16) showed the strongest negative response. Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significantly positively asspcoated, while species which were cooperative breeders (z = 2.53, p = 0.011), used grass/shrubland habitats (z = 7.64, p < 0.001), and had no known nest aggregation pattern (z = 4.20, p < 0.001) were statistically negatively associated.

## Differences in non-phylogenetic and phylogenetic models
Non-phylogenetic and phylogenetic models had similar relative ranks of standardized trait estimates and statistical significance (Fig. 4; Appendix 5). However, residual brain size was not significant in either non-phylogenetic model approaches (Fig. 4), but significantly associated with urbanization in our phylogenetic modelling, showing the strongest relative response in the phylogenetic global model (Fig. 4; Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important in the phylogenetic models, compared with the non-phylogenetic models (Fig. 5; Appendix 5). Lastly, body size was minimally important for the non-phylogenetic models, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5).


# DISCUSSION
The most urban-tolerant species in our analyses, defined based on their urbanization indices, had large niche breadths, reflected in generalism. Conversely, more specialized species were the least urban-tolerant: specialized feeding strategies (i.e., granivorous and insectivorous diets) seldom occurred in urban environments. These results were robust across separate non-phylogenetic and phylogenetic models, model-averaging approaches, and across phylogenetic uncertainty (Fig. S4), indicating that urbanization is generally excluding specialist species [@sorace2009distribution; @concepcion2015impacts]. Further, the non-random distribution of the urbanization index across the phylogeny (Fig. 3 & Table 2) is consistent with reduced phylogenetic diversity in urban areas [@la2018phylogenetic]. Particular phylogenetic effects [@derrickson1988taxon; @mckitrick1993phylogenetic] probably allow specific subsets of species to remain tolerant of urban environments [@morelli2016evidence; @ibanez2017global; @sol2017urbanisation].

This strong phylogenetic signal contradicts other studies [@evans2011makes], but was unsurprising given our large sample size of 477 species compared to the next highest of 55 species [@evans2011makes]. We also demonstrated the importance of treating species as non-independent, by considering phylogeny in the context of the linear models, highlighting the importance of investigating trait-urbanization relationships in a phylogenetically informed manner. For example, some of the more important traits in non-phylogenetic models (e.g., clutch size and feeding habitat generalism; Figures 4 and 5) had a stronger relationship (i.e., increased parameter estimates) with urbanization in the phylogenetic models than non-phylogenetic models. Other traits showed the opposite trend. Most notably, residual brain mass (and body size to a lesser extent) was significantly important within the phylogenetic models, but not in the non-phylogenetic models. This indicates that within clades, species with the largest residual brain mass [@maklakov2011brains] and largest body size [@major2010museum] were most suited to urban-tolerance. Some have indicated poor evidence that urban-tolerant species have large residual brain size [@kark2007living; @evans2011makes], but we found it was highly correlated with urban-tolerance [Fig. 4; Fig. 5; @moller2009successful; @maklakov2011brains; @moller2015brain].

Although many traits can identify a species' adaptability to urban environments, the published evidence has remained equivocal. Previously, much of the current focus was on discrete classifications of species, based on their responses to urbanization [@geschkecompact]. Contrastingly, we assessed the applicability of traits relative to one another; we did not run separate models or statistical tests for each specific trait [e.g., @kark2007living], allowing comparable effect sizes. There appear to be some consistent and generalizable trends within the trait-urbanization complex across the globe. For instance, our identification of the relative importance of annual fecundity was consistent with work from Europe [@croci2008does; @moller2009successful]. Urbanization also appears to select against insectivorous species [@chace2006urban; @evans2011makes]. The degree of sociality was also generally positively associated with urbanization, as found elsewhere [@coleman1994neophobia; @jokimaki1998distribution; @kark2007living]. However, our study also highlighted differences with global norms. For our data, urbanization selected against granivorous species, not previously identified [cf. @chace2006urban; @kark2007living; @evans2011makes]; this could be an artefact of relatively little supplementary feeding in Australian cities [@jones2018birdfeeding] compared to European cities [@fuller2008garden].

The large scale of our analysis was only possible following successful integration of broad-scale empirical datasets (i.e., citizen science data [@silvertown2009new] and remote sensing technology [@gorelick2017google]), which aid our understanding of the impacts of urbanization on birds [e.g., @aronson2014global; @aronson2016hierarchical; @gutierrez2018citizen]. We did not explicitly incorporate temporal changes in relation to the life cycle for different species' [e.g., @la2014role; @lepczyk2017global; @katuwaleffect], but migratory status was considered. In future, especially within migrant-dominated systems (e.g., Europe, North America), temporal changes in association with our urbanization index would add considerable value [@hostetler2015full; @marra2015call]. Australia has only a recent history of rapid urbanization [@coffee2016visualising] and our results may differ for parts of the world with long histories of development, and at different spatial scales [@suarez2018scale]. Yet, there is an opportunity to test our model predictions in urban environments in other parts of the world, potentially allowing prediction of which species face the greatest risk of impending urbanization.

Understanding which traits promote urban-tolerance or intolerance can inform actions to maintain avian biodiversity in urban environments [@evans2011makes]. In particular, future urban development could maintain biodiversity through improved protection of intact native vegetation [@sushinsky2013should] connected by corridors [@litteral2017role], advantaging specialist species detrimentally affected by landscape fragmentation and disturbance [@devictor2008distribution]. Further, increasing insect abundance and diversity in urban environments is potentially important [@baldock2015uk], given collapsing insect populations around the world [@hallmann2017more]. A species-specific approach could help to enhance urban biodiversity, which assesses foraging [@shochat2004linking] or nesting [@fernandez2001habitat] requirements of each species, for example, used to guide protection, mitigation, and restoration of necessary habitats in cities. Our urbanization index could provide quantifiable targets for local restoration projects, where a project aims to reduce the mean urbanization index for their local avifauna.

In conclusion, our novel continuous urbanization index showed relationships between ecological and life history traits with avian tolerance to urban environments. Improving understanding of the temporal nature of these relationships will be important, but we highlight the importance of considering these questions using a continuous measure of urbanization [@evans2011makes] as well as using a phylogenetically informed approach. It is the generalist species [@shochat2006patterns], with high clutch sizes and large residual brain sizes, which are best prepared to forage and survive in the ecologically novel aspects of urban environments.


# Data, code and materials
This work relied heavily on open-source, fully accessible datasets.
The necessary portions of those datasets, along with code to reproduce the analyses are available at
this github repository (https://github.com/cornwell-lab-unsw/bird_urbanness), and will be
permanently archived upon acceptance of the manuscript. The supporting data has been included in the article's supplementary material (Appendix 4).

# Competing interests
We have no competing interests.

# Author's Contributions
CTC, WKC, and JHW conceived and designed the methodology of the
study. CTC, WKC performed the statistical analyses. JMM, REM, and RTK contributed in drafting
the manuscript. All authors provided editorial advice, approved the final version of the manuscript,
and agree to be accountable for all aspects of the work.

# Acknowledgments
We thank the > 6,000 eBirders who have contributed data in Australia. We also thank the dedicated eBird team at the Cornell Lab of Ornithology and the global network of reviewers. Thanks to Stephen Garnett and Glenn Ehmke for providing us with extent of occurrence estimates for each of the species in our analysis. Thanks to Corrine Edwards for providing the illustrations used in Fig. 3.

# Funding
The authors declare no sources of funding specific to this work.

\newpage
# REFERENCES

<div id="refs"></div>

\newpage

# Figures

```{r Figure 1, echo=FALSE, fig.cap="Average radiance density distributions, on a logarithmic scale, across the range of five species with different distributional responses. The urbanization index, used as the response variable in our analyses, was taken as the median for a given species' distribution."}
knitr::include_graphics("finalFigs/Figure1.pdf")
```


```{r Figure 2, echo=FALSE, fig.cap="Histogram showing the count of the 477 species, relative to the urbanization index, on a logarithmic scale, highlighting the five species in Fig. 1. (See Figure S1 for randomly chosen indices for a given point to help orient the urbanization index.)"}
knitr::include_graphics("finalFigs/Figure2.pdf")
```


```{r Figure 3, echo=FALSE, fig.cap="Reproduced phylogenetic tree (Jetz et al. 2012) of our 477 species (see Appendix 4 for a full list of species with their urbanization indices), demonstrating clustering of species' associations with the urbanization index relative to phylogeny. There are two representations of the urbanization index for each species: (1) coloring ranges from a low urbanization index (dark red) to a high urbanization index (dark blue), with (2) magnitude of the urbanization index (barplot). The urbanization index was the median of the distribution for a species' average radiance values, as shown in Fig. 1. For reference, Figure S3 is a phylogenetic tree with the names of each tip displayed, and we also provide an interactive visualization (https://johnwilshire.github.io/bird_urbanness_web_app/)."}
knitr::include_graphics("finalFigs/Figure3.pdf")
```


```{r Figure 4, echo=FALSE, fig.cap="Parameter estimates for each of the four modelling approaches employed (See Appendix 5 for full model results), where a star represent significant effects. Model-averaged results did not contain every possible trait variable."}
knitr::include_graphics("finalFigs/Figure4.pdf")
```


```{r Figure 5, echo=FALSE, fig.cap="Proportional differences for each of the 22 traits' (Table 1) standardized parameter estimates, for the 477 species. Proportional differences represented the difference between a given trait's parameter estimate for the non-phylogenetically controlled and phylogenetically controlled global models."}
knitr::include_graphics("finalFigs/Figure5.pdf")
```


\newpage

# Tables
\small
```{r Table 2, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
library(dplyr)
Table1a <- read_csv("Data/Table1.csv")
colnames(Table1a) <- gsub("X1", "Variable", colnames(Table1a))
is.num <- sapply(Table1a, is.numeric)
Table1a[is.num] <- lapply(Table1a[is.num], round, 3)
Table1a <- Table1a %>%
  rename(`Abouheif's Cmean` = Cmean) %>%
  rename(`Moran's I` = I) %>%
  rename(`Blomberg's K` = K) %>%
  rename(`K*` = K.star) %>%
  rename(`Pagel's Lambda` = Lambda)

Table1b <- read_csv("Data/TableS2.csv")
colnames(Table1b) <- gsub("X1", "Variable", colnames(Table1b))
is.num <- sapply(Table1b, is.numeric)
Table1b[is.num] <- lapply(Table1b[is.num], round, 3)
Table1b <- Table1b %>%
  rename(`Abouheif's Cmean` = Cmean) %>%
  rename(`Moran's I` = I) %>%
  rename(`Blomberg's K` = K) %>%
  rename(`K*` = K.star) %>%
  rename(`Pagel's Lambda` = Lambda)

grouped <- rbind(cbind(Value="Statistic", Table1a), cbind(Value="p-values", Table1b))
tgrp <- xtabs(cbind(`Abouheif's Cmean`, `Moran's I`, `Blomberg's K`, `K*`, `Pagel's Lambda`) ~ Value+Variable, grouped)

table <- ftable(aperm(tgrp, c(3, 1, 2)), row.vars=c("Variable", "Value"))

library(kableExtra)
library(pander)
pander(table, caption = "Summary of results showing an assessment of a phylogenetic signal for our response variable (urbanization index) and the continuous predictor traits. We show all available indices (Keck et al. 2016), each of which represents the statistical nonindependence of a trait's value due to its phylogenetic relatedness (Revell et al. 2008).", split.table=Inf)

```

\newpage

# Supplementary figures
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }
     \beginsupplement

```{r Figure S1, echo=FALSE, fig.cap="The VIIRS nightime-lights shown for Australia, Brisbane (top right), Melbourne (bottom left), and Sydney (bottom right). Average radiance of VIIRS night-time lights, in red, was calculated for a given checklist within a 5 km buffer, to match the spatial scale of checklists of Australia's bird species. Values shown demonstrate the average radiance value for seven different points in Australia."}
knitr::include_graphics("finalFigs/FigureS1.pdf")
```

```{r Figure S2, echo=FALSE, fig.cap="Correlations between continuous traits, number of records, and number of unique localities used in the analysis."}
knitr::include_graphics("finalFigs/FigureS2.pdf")
```

```{r Figure S3, echo=FALSE, fig.cap="Reference phylogenetic tree of the 477 bird species in the analysis, with labeled tips, based on Jetz et al. (2012). We also provide an interactive visualization for better interpreting the phylogenetic tree (https://johnwilshire.github.io/bird_urbanness_web_app/)"}
knitr::include_graphics("finalFigs/FigureS3.pdf")
```

```{r Figure S4, echo=FALSE, fig.cap="The same phylolm model was fitted to 1000 of the family of trees presented by Jetz et al. (2012) to account for one aspect of phylogenetic uncertainty. Density plots show estimated parameters for the most important traits across models."}
knitr::include_graphics("finalFigs/FigureS4.pdf")
```

\newpage    

# Supplementary tables
```{r Table S1, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)
library(dplyr)

collinearity_investigation <- read_csv("Data/TableS1.csv")
collinearity_investigation <- collinearity_investigation[-1,]
collinearity_investigation <- collinearity_investigation %>%
  rename(Trait = X1)
collinearity_investigation <- collinearity_investigation[c(2, 1, 4, 16, 3, 6, 11, 5, 22, 20, 15, 14, 10, 7, 8, 9, 13, 12, 21, 18, 19, 17), ]

kable(collinearity_investigation, format = "latex", caption = "Variance inflation factors for the predictor traits used in the analysis.", booktabs = T) %>%
kable_styling(latex_options = c("hold_position"))

```
